---
title: "Final_ARM"
author: "Alfred Nick Tinsley"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Code for final Project



### Setting working directory for my data sets
```{r wd}

setwd("C:/Users/nicktinsley/OneDrive/Syracuse/R code/IST 707/data/finalProject")

```



### Loading Libraries for our discovery
```{r libraries}
library(arules)
library(arulesViz)
library(plyr)
library(dplyr)
library(funModeling)

```


### Load data sets 


```{r red_wht}

red <- read.csv2("red.csv")
white <- read.csv2("white.csv")

```


### Viewing Data

```{r }

View(red[red$quality > 5,])

hist(red$quality)
hist(white$quality)

```

### Discretize Categories
##quality 5 - 7 mid
##quality < 5 low
##quality > 7 high


### Munge data set
1. Convert char to factor


```{r}
red <- red %>% mutate_if(is.character, funs(as.factor))
str(red)

white <- white %>% mutate_if(is.character, funs(as.factor))
str(white)


```


2. Discretize the quality attribute

```{r}
red$quality_disc <- cut(red$quality, breaks = c(0,4,6,Inf),
                        labels=c("low","mid","high"))

table(red$quality, red$quality_disc)

white$quality_disc <- cut(white$quality, breaks = c(0,4,6,Inf),
                        labels=c("low","mid","high"))

table(white$quality, white$quality_disc)





```

3 . Remove quality attribute so it does not create unecesarry rules

```{r}
red <- red %>% select(-quality)
white <- white %>% select(-quality)

```


4. Create one data set for all wines to see if rules for each compare with combined DS

```{r}
red_white <- rbind(red,white)

```

### Create some rules for discovery

* Red Wine Rules
```{r red_wine}
rules_highQ <- apriori(data=red, parameter = list(supp=0.001, conf=0.08,minlen=2),
                       appearance = list(default="rhs",lhs="quality_disc=high"),
                       control=list(verbose=T))
rules_highQ <- sort(rules_highQ, decreasing = T, by = "lift")
inspect(rules_highQ)

#mid 
rules_midQ <- apriori(data=red, parameter = list(supp=0.001, conf=0.08,minlen=2),
                       appearance = list(default="rhs",lhs="quality_disc=mid"),
                       control=list(verbose=T))
rules_midQ <- sort(rules_midQ, decreasing = T, by = "lift")
inspect(rules_midQ)

#low
rules_lowQ <- apriori(data=red, parameter = list(supp=0.001, conf=0.08,minlen=2),
                      appearance = list(default="rhs",lhs="quality_disc=low"),
                      control=list(verbose=T))
rules_lowQ <- sort(rules_lowQ, decreasing = T, by = "lift")
inspect(rules_lowQ)



```




* White Wine Rules

```{r white_wine}
rules_wht <- apriori(white, parameter=list(supp=0.001, conf=0.9, maxlen=4))
rules_wht <- sort(rules_wht, decreasing = TRUE, by ="confidence")
inspect(rules_wht[1:50])



```




* Combined Red White Rules

```{r red_wht_wine}
rules_redwht <- apriori(red_white, parameter=list(supp=0.001, conf=0.9, maxlen=4))
rules_redwht <- sort(rules_redwht, decreasing = T, by="confidence")
inspect(rules_redwht[1:50])


rules_rdwht_mid <- apriori(data=red_white, parameter = list(supp=0.001, conf=0.9,maxlen=4),
                           appearance = list(default="lhs",rhs="quality_disc=mid"),
                           control=list(verbose=T))
rules_rdwht_mid <- sort(rules_rdwht_mid, decreasing = T, by="lift")
inspect(rules_rdwht_mid[1:10])

rules_rdwht_high <- apriori(data=red_white, parameter = list(supp=0.001, conf=0.9,maxlen=4),
                            appearance = list(default="lhs",rhs="quality_disc=high"),
                            control=list(verbose=T))
rules_rdwht_high <- sort(rules_rdwht_high, decreasing = T, by="lift")
inspect(rules_rdwht_high[1:10])


```



















